Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-01-06 Manuscript/TEM_draft.tex
Updated intro/methods/results
Results only for the table

\subsubsection{Summarizing the comparisons}
Finally after doing all theses %stupid
50 simulations and comparisons within, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework.
% TG: the following part might be useless
To
To do this we measured the similarity among the different distributions % NC: OF WHAT!!! tree metrics for all of the analyses above
using the Bhattacharyya Coefficient \citep{Bhattacharyya}. The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} % FOR WHAT!!!).

%For each method (Maximum Likelihood trees, Bayesian %MCC?
%trees, Bootstraps and Bayesian posterior trees) and metric (Robinson-Foulds and Triplets distance), we used this coefficient in two ways in order to:
\begin{enumerate}
\item{Assess the effect of the method on all parameters:}

We performed a pairwise comparison of each parameter between each pair of methods. This resulted in 125 comparisons per method and per metric (each parameter combination within one method compared to the same parameter combination in the other method). We then compared the distribution of these pairwise comparisons to see the global difference between two methods: if the method have really similar results, we would expect the distribution to be cluster around 1 (high probability of overlap between the distributions) and if the methods are really dissimilar around 0 (low probability of overlap - Fig.~\ref{Fig_Results-paircomp_without}).
\item{Assess the effect of the parameters within a method:}

We performed a pairwise comparison per method and per metric for each combination of parameters. This resulted in 7875 pairwise comparisons per method and per metric (triangle of a 125$\times$125 matrix). We represented these results as a triangular matrix with the values of each pairwise comparison coloured according to the value of the Batthacharyya Coefficient (coloured in green when the distributions overlap completely and in red when they don't overlap at all- Fig.~\ref{Fig_Results-paircomp_within}).
\end{enumerate}

% Add in your three questions here?
%---------------------------------------------
%
%       RESULTS
%
%---------------------------------------------

@@

>>>>

\subsubsection{Summarizing the comparisons}
Finally after doing all theses %stupid
50 simulations and comparisons within, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework. Additionally, we also assessed the effect of the different methods (i.e. Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).% and the joint effect of the our missing data parameters and the methods.
% TG: the following part might be useless (complete repetition of above)
To assess the ability to recover the "correct" tree topology on our 50 simulations, we used the scores for both normalised metrics (Robinson-Foulds and Triplets distance). We looked at the following effects:
\begin{enumerate}
\item{The three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$);}
\item{The methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
%\item{The joint effect of the missing data parameters and the methods.} actually this is implied in both already.
\end{enumerate}
% TG: end of the repetition

%Introducing Bhattacharyya Coefficient
To evaluate these effects we measured the similarities among the different normalised metrics scores (Robinson-Foulds and Triplets distance) for all the analysis above using the Bhattacharyya Coefficient \citep{Bhattacharyya}. The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).

%Explaining the parameter pairwise comparisons (ANOVA equivalent=within)
For assessing the effect of our three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$), we calculated the Bhattacharyya Coefficient between each pairwise combination for each parameter state (0\%, 10\%, 25\%, 50\%, 75\%). This resulted in 7875 pairwise distributions pairwise comparisons (triangular $3^5$$\times$$3^5$ matrix). We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient=1) and in red when they don't (Bhattacharyya Coefficient=0) - Fig.~\ref{Fig_Results-paircomp_within}). We repeated this procedure for all the methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).

%Explaining the methods comparisons (ANOVA equivalent=among)
Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g. Maximum Likelihood\textit{versus} Bayesian consensus). We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e. the Bhattacharyya Coefficients) in table ~\ref{Tab_Results-Difference_methods}. When two methods display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees, we expect the Bhattacharyya Coefficients to be close to 1. On the opposite, if the two methods displays dissimilar normalised metrics scores, we expect the Bhattacharyya Coefficients to be close to 0.

%---------------------------------------------
%
%       RESULTS
%
%---------------------------------------------

@@

\begin{table}[ht]
\label{Tab_Results-Difference_methods}
\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data".}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
    Bayesian consensus              & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
    Bootstraps                      & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
    bootstraps                      & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
    Boostraps \textit{vs.}          & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
    Bayesian posterior tree         & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}
